---
title: "Professor Oak's thesis"
author: "Deon Roos & Ross Kwok & Tom Price"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    highlight: monochrome
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
# Packages: --------------------------------------------------------------------
library(googlesheets4) # For loading in the data from google sheets
library(ggplot2)       # For data visualiations
library(lme4)          # For mixed effect linear models
library(mgcv)          # For non-linear models
library(ggeffects)     # For quick model figures
library(tidyverse)     # For data manipulation
library(plotly)        # For interactive plots
theme_set(theme_minimal())
```

# Pokemon Go analysis

Just a document to mess around with some data and `.rmd` documents.

```{r}
# Load the data from googlesheets (requires authorisation): --------------------
df <- read_sheet("https://docs.google.com/spreadsheets/d/1EWzGk_qDK8ommXYz2jxYvFSSEzj9Wal976dWRwR4_0w/edit?usp=sharing",
                 #24, 
                 sheet = "Data", 
                 trim_ws = TRUE)

# Minor data tidy up: ----------------------------------------------------------
# Treat following as factors
df$player <- factor(df$player)
df$num_evo <- factor(df$number_evolutions) # Could be treated as number
df$patch <- factor(df$patch)
df$pokemon <- factor(df$pokemon)
df$type <- factor(df$type)
df$type_2 <- factor(df$type_2)
# Log difference growth rate in cp
df$cp_lambda <- log(df$final_cp) - log(df$starting_cp)
# Changing text shorthand for full word
df$evolve_stone <- factor(ifelse(df$evolve_stone == "y", "Yes", "No"))
df$special <- factor(ifelse(df$special == "y", "Yes", "No"))
```


# Summary figures

```{r}
df2 <- df |> 
  group_by(pokemon) |> 
  summarise(pokemon_freq = n())
ggplot(df2[df2$pokemon_freq > 10,], aes(x = reorder(pokemon, pokemon_freq), y = pokemon_freq, fill = pokemon)) +
  geom_col(show.legend = FALSE) +
  labs(x = "Pokemon",
       y = "Number of data entries",
       title = "Most common pokemon in the data") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1))
```

```{r}
df2 <- df |> 
  group_by(pokemon) |> 
  summarise(mean_start = mean(starting_cp))
ggplot(df2[df2$mean_start > 1000,], aes(x = reorder(pokemon, mean_start), y = mean_start)) +
  geom_point() +
  labs(x = "Pokemon",
       y = "Mean starting CP",
       title = "Pokemon with highest average starting CP") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1))
```

```{r}
ggplot(df2[df2$mean_start < 300,], aes(x = reorder(pokemon, mean_start), y = mean_start)) +
  geom_point() +
  labs(x = "Pokemon",
       y = "Mean starting CP",
       title = "Pokemon with lowest average starting CP") +
  theme(axis.text.x = element_text(angle = 90, 
                                   vjust = 0.5, 
                                   hjust = 1))
```

```{r}
ggplotly(ggplot(df, aes(x = starting_cp, y = final_cp, colour = cp_diff, text = pokemon)) +
           geom_point() +
           labs(x = "Starting CP",
                y = "Final CP",
                colour = "Change in CP",
                title = "Change in pokemon CP after evolving") +
           scale_colour_viridis_c(option = "C")
)
```

```{r}
ggplotly(ggplot(df, aes(x = starting_cp, y = final_cp, colour = cost_evolve, text = pokemon)) +
           geom_point() +
           labs(x = "Starting CP",
                y = "Final CP",
                colour = "Cost to evolve",
                title = "Pokemon with lowest average starting CP") +
           scale_colour_viridis_c(option = "C")
)
```

```{r}
ggplotly(ggplot(df, aes(x = starting_cp, y = final_cp, colour = player_level, text = pokemon)) +
           geom_point() +
           labs(x = "Starting CP",
                y = "Final CP",
                colour = "Player level",
                title = "Pokemon with lowest average starting CP") +
           scale_colour_viridis_c(option = "C")
)
```

# Analysis

Above figures suggest strong linear relationship between starting CP and final CP but with interacting effects of player level and cost to evolve. Will include random effects for pokemon (to account for varied effort and partially pool pokemon together), primary type, secondary type. Will need to be log-Normal model so that final CP is forced to be positive.

```{r}
m1 <- lmer(final_cp ~ starting_cp * player_level * cost_evolve + 
             (1 | pokemon),
           data = df)
summary(m1)
plot(ggpredict(m1))
```

```{r}
nu_data <- expand.grid(
  starting_cp = seq(from = min(df$starting_cp), to = max(df$starting_cp), length.out = 20),
  player_level = seq(from = min(df$player_level), to = max(df$player_level), length.out = 20),
  cost_evolve = seq(from = min(df$cost_evolve), to = max(df$cost_evolve), length.out = 20),
  pokemon = "Charmander"
)

nu_data$pred <- predict(m1, nu_data)

ind <- exclude.too.far(
  nu_data$starting_cp, nu_data$player_level,
  df$starting_cp, df$player_level, 
  0.1)

nu_data[ind,] <- NA

ggplot() +
  geom_tile(data = nu_data, aes(x = starting_cp, y = player_level, fill = pred)) +
  scale_fill_viridis_c(option = "C") +
    geom_point(data = df, aes(x = starting_cp, y = player_level),
             size = 0.1, alpha = 0.2) +
  labs(x = "Starting CP",
       y = "Player level",
       fill = "Predicted\nfinal CP")
```

```{r}
nu_data <- expand.grid(
  starting_cp = seq(from = min(df$starting_cp), to = max(df$starting_cp), length.out = 20),
  player_level = seq(from = min(df$player_level), to = max(df$player_level), length.out = 20),
  cost_evolve = seq(from = min(df$cost_evolve), to = max(df$cost_evolve), length.out = 20),
  pokemon = "Charmander"
)

nu_data$pred <- predict(m1, nu_data)

ind <- exclude.too.far(
  nu_data$starting_cp, nu_data$cost_evolve,
  df$starting_cp, df$cost_evolve, 
  0.1)

nu_data[ind,] <- NA

ggplot() +
  geom_tile(data = nu_data, aes(x = starting_cp, y = cost_evolve, fill = pred)) +
  scale_fill_viridis_c(option = "C") +
  geom_point(data = df, aes(x = starting_cp, y = cost_evolve),
             size = 0.1, alpha = 0.2) +
  labs(x = "Starting CP",
       y = "Cost to evolve",
       fill = "Predicted\nfinal CP")
```

```{r}
nu_data <- expand.grid(
  starting_cp = seq(from = min(df$starting_cp), to = max(df$starting_cp), length.out = 20),
  player_level = seq(from = min(df$player_level), to = max(df$player_level), length.out = 20),
  cost_evolve = seq(from = min(df$cost_evolve), to = max(df$cost_evolve), length.out = 20),
  pokemon = "Charmander"
)

nu_data$pred <- predict(m1, nu_data)

ind <- exclude.too.far(
  nu_data$player_level, nu_data$cost_evolve,
  df$player_level, df$cost_evolve, 
  0.05)

nu_data[ind,] <- NA

ggplot() +
  geom_tile(data = nu_data, aes(x = player_level, y = cost_evolve, fill = pred)) +
  scale_fill_viridis_c(option = "C") +
  geom_point(data = df, aes(x = player_level, y = cost_evolve),
             size = 0.1, alpha = 0.2, colour = "black") +
  labs(x = "Player level",
       y = "Cost to evolve",
       fill = "Predicted\nfinal CP")
```

```{r}
nu_data <- expand.grid(
  starting_cp = seq(from = min(df$starting_cp), to = max(df$starting_cp), length.out = 20),
  player_level = seq(from = min(df$player_level), to = max(df$player_level), length.out = 20),
  cost_evolve = c(25, 50, 100, 400),
  pokemon = "Charmander"
)

nu_data$pred <- predict(m1, nu_data)

ind <- exclude.too.far(
  nu_data$player_level, nu_data$starting_cp,
  df$player_level, df$starting_cp, 
  0.04)

nu_data[ind,] <- NA

nu_data$cost_evolve <- factor(nu_data$cost_evolve)

nu_data <- nu_data[complete.cases(nu_data),]

ggplot() +
  geom_tile(data = nu_data, aes(x = starting_cp, y = player_level, fill = pred)) +
  scale_fill_viridis_c(option = "C") +
  facet_wrap(~cost_evolve, scales = "free") +
  labs(x = "Starting CP",
       y = "Player level",
       fill = "Predicted\nchange in CP")
```

